---
title: "Assignment 4 - Aquaculture"
author: "Gabrielle"
date: '2022-11-09'
output: 
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,message=FALSE,warning = FALSE,results=FALSE)
library(tidyverse)
library(here)
library(ggpubr) # for some graphic applications that extend ggplot2
library(janitor)
library(broom) # used to make tables
library(knitr) # used to make table
```

Research Question: Does Environmental DNA metabarcoding estimate the species area relationship (SAR) the same or different than traditional fishing gear?

Suggestions and hints:
1. You may need to transform your data or use a log function in the model formulation (see SAR formulation in the lecture notes)
2. Make sure you address the assumptions of all analyses and make caveats about any violations of the assumptions
3. Be concise, yet thorough. Points will be deducted for excessive figures or analyses irrelevant to your argument.
4. You will need to bring the data into tidy form for analyses and plots.

##### Load in the data + clean it:

- area_ha: The area of the lake surveyed in hectares
- dna_richness: The number of unique fish species detected using eDNA metabarcoding
- trad_richness: The number of unique fish species detected using traditional/conventional fishing approaches.

```{r}
sar <- read_csv(here("data","SAR_data.csv"))

# use janitor() and clean up the names
sar <- sar %>% clean_names()
```

##### Scatterplot of the data to visually inpect it: dna_richness

```{r sar_1, fig.align='center',fig.cap="Figure 1. Scatterplot of the relationship between the area of the lake surveyed (in hectares) and the number of unique fish species detected using eDNA metabarcoding."}

sar_scatter_dna <- ggplot(sar, aes(x = area_ha, y = dna_richness)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(x = "Area of lake surveyed (hectares)", y = "# of unique fish detected via eDNA") +
  theme_minimal()

sar_scatter_dna
```

##### Scatterplot of the data to visually inpect it: trad_richness

```{r sar_2, fig.align='center',fig.cap="Figure 2. Scatterplot of the relationship between the area of the lake surveyed (in hectares) and the number of unique fish species detected using traditional/conventional fishing approaches."}

sar_scatter_trad <- ggplot(sar, aes(x = area_ha, y = trad_richness)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(x = "Area of Lake Surveyed (hectares)", y = "# of Unique Fish Detected via Traditional Approaches") +
  theme_minimal()

sar_scatter_trad
```


##### Log transform the data:

```{r}
sar$area_ha_mod = log(sar$area_ha)

sar$dna_richness_mod = log(sar$dna_richness)

sar$trad_richness_mod = log(sar$trad_richness)

```


Re-visualize using the log-transformed data:

```{r sar_3, fig.align='center',fig.cap="Figure 3. Scatterplot of the relationship between the logged area of the lake surveyed (in hectares) and the logged number of unique fish species detected using eDNA metabarcoding."}

sar_scatter_dna_mod <- ggplot(sar, aes(x = area_ha_mod, y = dna_richness_mod)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(x = "Logged Area of Lake Surveyed (hectares)", y = "Logged # of Unique Fish Detected via eDNA") +
  theme_minimal()

sar_scatter_dna_mod
```

```{r sar_4, fig.align='center',fig.cap="Figure 4. Scatterplot of the relationship between the logged area of the lake surveyed (in hectares) and the logged number of unique fish species detected using traditional/conventional fishing approaches."}

sar_scatter_trad_mod <- ggplot(sar, aes(x = area_ha_mod, y = trad_richness_mod)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(x = "Logged Area of Lake Surveyed (hectares)", y = "Logged # of Unique Fish Detected via Traditional Approaches") +
  theme_minimal()

sar_scatter_trad_mod
```


